Combining formal and distributional models of predicate semantics

### 2016-03-07 papers/acl2016/paper.tex
A bunch of work on the paper
Former-commit-id: 3c90e1d56ce2c57dd47a9f0ba5c248f0a283df78

In this section we present our models of predicate semantics.  In
\secref{formal-and-distributional} we describe how we modify Krishnamurthy and
Mitchell's (K&M's) semantic parser to include features from a formal knowledge base.
In \secref{computing-pmi} we describe how we select those features for each
predicate.  The remaining two sections then describe other improvements we made
to K&M's semantic parsing model: lexicalizing preposition and noun-compound
predicates, and using the Freebase graph to select candidate entities.

@@

The modification we make to K&M's semantic parser is simple.  For each entity
and entity pair in the data we compute an SFE feature vector $\psi(e)$ and
$\psi(e_1, e_2)$.  For each category and relation, we select a subset of the
features to have associated weights $\omega_c$ and $\omega_r$.  We then add the
weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

@@

Note that now there are three sets of parameters to be learned by the semantic
parser: (1) $\theta$, low-dimensional distributional vectors trained for each
predicate; (2) $\phi$, low-dimensional distributional vectors trained for each
entity and entity pair; and (3) $\omega$, weights associated with the selected
formal SFE features for each predicate.  All of these parameters are optimized
jointly, using the same ranking objective used by K&M.

@@

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training

@@

@@

@@

>>>>

In this section we present our models of predicate semantics.  In
\secref{formal-and-distributional} we describe how we modify Krishnamurthy and
Mitchell's (K\&M's) semantic parser to include features from a formal knowledge base.
In \secref{computing-pmi} we describe how we select those features for each
predicate.  The remaining two sections then describe other improvements we made
to K\&M's semantic parsing model: lexicalizing preposition and noun-compound
predicates, and using the Freebase graph to select candidate entities.

@@

The modification we make to K\&M's semantic parser is simple.  For each entity
and entity pair in the data we compute an SFE feature vector $\psi(e)$ and
$\psi(e_1, e_2)$.  For each category and relation, we select a subset of the
features to have associated weights $\omega_c$ and $\omega_r$.  We then add the
weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

@@

Note that now there are three sets of parameters to be learned by the semantic
parser: (1) $\theta$, low-dimensional distributional vectors trained for each
predicate; (2) $\phi$, low-dimensional distributional vectors trained for each
entity and entity pair; and (3) $\omega$, weights associated with the selected
formal SFE features for each predicate.  All of these parameters are optimized
jointly, using the same ranking objective used by K\&M.

@@

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training data to find a mapping from
entities and entity pairs to predicates that they have been seen with in the
data.  For example, the phrase ``Italian architect Andrea Palladio'' is
considered a positive training example for the instantiated predicates
$\predicate{architect}(\entity{Andrea Palladio})$ and
$\predicate{architect\_N/N}(\entity{Italy}, \entity{Andrea Palladio})$.  We
then associate the feature vectors for \entity{Andrea Palladio} and
(\entity{Italy}, \entity{Andrea Palladio}) with the predicates
\predicate{architect} and \predicate{architect\_N/N}, respectively.  For each
predicate $\pi$ and feature $f$, we use these counts to calculate the
probabilities $p(\pi, f)$, $p(\pi)$ and $p(f)$.  After removing features with
counts below some threshold, we pick the $k$ features with the highest PMI
values for each predicate to use in our model.

@@

While we largely follow the semantic parsing framework introduced by K\&M, we
made one important change in how logical forms are generated.  For the running
example we have been using this paper, ``Italian architect Andrea Palladio'',
the relation predicate produced by K\&M's system is
$\predicate{N/N}(\entity{Italy}, \entity{Andrea Palladio})$.  \predicate{N/N}
does correctly capture the fact that \entity{Italy} is used syntactically as a
noun modifier of \entity{Andrea Palladio}, but it leaves out the additional
information contained in the syntax that ``architect'' was mediating that
noun-noun dependency.  This means that ``U.S. president Barack Obama'' and
``Italian architect Andrea Palladio'' are both modeled using the same relation
predicate, which leads to a very generic and largely useless model for the
predicate \predicate{N/N}.  By splitting \predicate{N/N} into separate
predicates (\predicate{architect\_N/N} and \predicate{president\_N/N}), we give
the model the opportunity to learn more fine-grained relation semantics, and we
allow our feature selection pipeline to find much more useful features for
relation predicates.  While it would be ideal to use actual dependency
relations to determine which words to add to the predicate in complex noun
phrases, most current parsers treat these kinds of noun phrases as flat, with
all noun modifiers depending directly on the head of the noun phrase.  We
approximate this ideal by simply taking all words in between the two entities
in a noun phrase as the predicate (e.g., ``Illinois attorney general Lisa
Madigan'' would produce the predicate \predicate{attorney\_general\_N/N}).

A similar issue arises with prepositions and possessives.  For a phrase such as
``Barack Obama, president of the U.S.'', K\&M's system would produce the
following relation predicate: $\predicate{of}(\entity{Barack Obama},
\entity{U.S.})$.  Because this same predicate (\predicate{of}) is used for all
instances of this preposition, the predicate is forced to model too much and
becomes overly generic.  When there is a common noun mediating a prepositional
dependency between two entities, as in the phrase above, we add that noun to
the predicate, giving the instance $\predicate{president\_of}(\entity{Barack
Obama}, \entity{U.S.})$.  We do the same for possessives in phrases such as
``Rome, Italy's capital'', producing the predicate \predicate{'s\_capital}.

@@

One final way in which we improved K\&M's semantic parser is in how candidate
entities are generated during inference.  Because the distributional-only
predicate models used by K\&M's semantic parser need to learn a vector of
parameters for each entity pair, they cannot give a non-zero score to any
entity pair not seen during training time.  When given a query such as
$\lambda(\entity{x}) \predicate{architect\_N/N}(\entity{x}, \entity{Andrea
Palladio})$, K\&M's system only needs to consider the set of entities
$\entity{x}$ that have been seen at training time with \entity{Andrea
Palladio}; all other entities will get a score of 0.

Our predicate models do not have this limitation.  The distributional component
will still give an unseen entity pair a score of 0, as it is identical to K\&M's
model, but the formal component of our simples simply computes an observed
feature vector from Freebase for the entity pair.  As long as both entities
appear in Freebase, our model can compute a feature vector and give a non-zero
score.

While this improvement allows us to handle queries over rare entities much
better than a distributional-only model, it also complicates inference, as we
cannot just score entity pairs seen at training time.  Optimizing every query's
score over all of the millions of entities in Freebase is not computationally
feasible.  To make inference tractable, for a given query we restrict
candidates to all entities seen with the query entities during training (as
done by K\&M), as well as all entities directly connected to the query entities
in Freebase, or connected by a mediator node\footnote{Mediators in Freebase are
used to capture relations with more than two arguments, such as employment
tenure, which has an employer, and employee, a start date, and an end date.}.

For some entities, such as the United States, the set of connected entities can
still be intractably large, so we limit this expansion to only those entities
with fewer than 100 directly connected entities.  Our motivation for this is
that this candidate entity generation is most useful for rarely seen entities,
for which we have few or no related entities seen during training.  These
entities also tend to have relatively few connected entities in Freebase.

@@

\mattnote{I will add more detail to some of these examples; I ran out of
time...}

Our model does very well on most predicates that have clear Freebase
correlates: prime ministers, presidents, newspapers, companies, and so on.  The
query ``French newspaper \blank{}'' gets almost perfect average precision.

It does less well on predicates that do not match as closely with Freebase,
such as ``Microsoft head honcho \blank{}'', ``Patriots linebacker \blank{}'',
etc.  The combined model handily beats the distributional-only model in most of
these cases, but it does not have the nearly-perfect AP seen in the more crisp
queries above.  The ``head honcho'' example is particularly informative---this
is a phrase that essentially means ``leader'', and the features learned by the
formal model get some essential hints from Freebase, even though there is no
direct correlate.  The features capture the fact that entities that are
``honchos'' are people, and that they tend to be CEOs and film producers.

For ``Patriots linebacker \blank{}'', the issue is that the features for
\predicate{linebacker\_N/N} are able to capture which athletes play for which
teams, but they cannot distinguish well between what positions those athletes
play, so Patriots quarterbacks, wide receivers, and running backs are scored
highly for this query.  This could be fixed by using more complex features for
relation predicates, instead of just edge sequences, but that would explode the
feature space and make the feature selection process more difficult.  It could
also be fixed by better negative example selection during training.

There are also cases where a predicate has so many possible entities types in
Freebase that the it cannot learn models that are useful for all queries.  A
good example of this is ``\blank{} is a NASA mission''.

The SFE features we used cannot capture time very well.  ``17th century French
mathematician \blank{}''...

Because of the simple logical forms produced by the semantic parser we used, we
cannot model some complex phenomena, such as counting and other things.  This
is an orthogonal issue to the predicate semantics we're concerned with in this
paper, however.

Vice president Al Gore is modeled as $\predicate{vice}(\entity{Al Gore}) \land
\predicate{president}(\entity{Al Gore})$.  Clearly there is room for better
processing of multi-word expressions.

\predicate{head\_honcho\_N/N} was never seen in the training data, so the
phrase ``Microsoft head honcho \blank{}'' has to rely on knowing whether an
entity is a ``honcho'' and what entities are related to Microsoft by an unknown
predicate.  \predicate{honcho\_N/N} \emph{is} seen in the training data,
however; some kind of backoff to simpler phrases would be useful here.

Using connected entities in Freebase as additional candidates during inference
is helpful, but it is not very sophisticated.  It over-generates for a lot of
entities and predicates, and under-generates for others.  A mechanism for
tailoring a search over Freebase to a particular query would be useful, though
that would move the model more into the realm of reinforcement learning.
